---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---


```{r,results='HIDE'}
library(ggplot2)
#library(scales)
#library(Hmisc)
```
Loading and preprocessing the data


1. Load the data (i.e. read.csv())
```{r }

activity_raw <- read.csv('activity.csv', header = TRUE, sep = ",",
                  colClasses=c("numeric", "character", "numeric"))

#Process/transform the data (if necessary) into a format suitable for analysis
# Transform the date attribute to an actual date format
activity_raw$date <- as.POSIXct(activity_raw$date, format="%Y-%m-%d")

# Compute the weekdays from the date attribute
activity_raw <- data.frame(date=activity_raw$date, 
                           weekday=tolower(weekdays(activity_raw$date)), 
                           steps=activity_raw$steps, 
                           interval=activity_raw$interval)

# Compute the day type (weekend or weekday)
activity_raw <- cbind(activity_raw, 
                      daytype=ifelse(activity_raw$weekday == "saturday" | 
                                     activity_raw$weekday == "sunday", "weekend", 
                                     "weekday"))

# Create the final data.frame
activity <- data.frame(date=activity_raw$date, 
                       weekday=activity_raw$weekday, 
                       daytype=activity_raw$daytype, 
                       interval=activity_raw$interval,
                       steps=activity_raw$steps)

# Clear the workspace
rm(activity_raw)
#We display the first few rows of the activity data frame:

head(activity)
```
What is the mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day

```{r}
sum_data <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=TRUE)
names(sum_data) <- c("date", "total")
hist(sum_data$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="green", 
     xlab="Total number of steps", 
     ylim=c(0, 20), 
     main="Histogram of the total number of steps taken each day")
   
mean(sum_data$total)
median(sum_data$total)
```

What is the average daily activity pattern?
```{r}
rm(sum_data)

# Compute the means of steps accross all days for each interval
mean_data <- aggregate(activity$steps, 
                       by=list(activity$interval), 
                       FUN=mean, 
                       na.rm=TRUE)

# Rename the attributes
names(mean_data) <- c("interval", "mean")
plot(mean_data$interval, 
     mean_data$mean, 
     type="l", 
     col="blue", 
     lwd=2, 
     xlab="Interval [minutes]", 
     ylab="Average number of steps", 
     main="Time-series of the average number of steps per intervals\n(NA removed)")
max_pos <- which(mean_data$mean == max(mean_data$mean))
max_interval <- mean_data[max_pos, 1]
```
The 5-minute interval that contains the maximum of steps, on average across all days, is `r max_interval`
Inputing the missing values
```{r}
rm(max_interval)

# We use the trick that a TRUE boolean value is equivalent to 1 and a FALSE to 0.
NA_count <- sum(is.na(activity$steps))
# Find the NA positions
na_pos <- which(is.na(activity$steps))

# Create a vector of means
mean_vec <- rep(mean(activity$steps, na.rm=TRUE), times=length(na_pos))
#Create a new dataset that is equal to the original dataset but with the missing data filled in.
# Replace the NAs by the means
activity[na_pos, "steps"] <- mean_vec

# Clear the workspace
rm(mean_vec, na_pos)
# Compute the total number of steps each day (NA values removed)
sum_data <- aggregate(activity$steps, by=list(activity$date), FUN=sum)

# Rename the attributes
names(sum_data) <- c("date", "total")

# Compute the histogram of the total number of steps each day
hist(sum_data$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="blue", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Histogram of the total number of steps taken each day\n(NA replaced by mean value)")
mean(sum_data$total)
median(sum_data$total)
```

Are there differences in activity patterns between weekdays and weekends?
```{r}

# Compute the average number of steps taken, averaged across all daytype variable
mean_data <- aggregate(activity$steps, 
                       by=list(activity$daytype, 
                               activity$weekday, activity$interval), mean)

# Rename the attributes
names(mean_data) <- c("daytype", "weekday", "interval", "mean")

ggplot(mean_data,aes(interval,mean))+ 
         geom_line(color="magenta") + 
        facet_wrap(~ daytype, nrow=2, ncol=1) +
        labs(x="5-minute Interval", y="Avg number of steps") +
        theme_bw()
    
```

